Chapter 16: Particle-laden gravity currents: the lock-release slumping regime at the laboratory scale

### 2024-03-03 paper/main.tex
new fig1

Other mechanisms can modify the dynamics of the slumping regime, possibly leading to other possible positions $X_{\text t}$ or times $T_{\text t}$ at which the velocity is no longer time-independent. Depending on the relative intensity of these mechanisms, this can occur prior to the influence of the rarefaction wave or bore. Hence, determining the various regimes related to the prevalence of each mechanism is part of the topic of the present chapter.

@@

A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics. Nevertheless, the prediction of these finite scales is difficult as it should strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation. In turn, this can modify the macroscopic behavior of the current, for instance, the dynamics of the front. Accordingly, these dynamics shall be controlled by several dimensionless parameters that can lead to a wide variety of flow regimes. In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $\mathcal{A}t$, the ratio between available inertia and dissipation, say a Reynolds number $\mathcal{R}e$ as usually introduced as a control parameter for any density currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity $S=V_{\text s}/U_{\text{c}}$ with $V_{\text s}$ the settling velocity of particles into the carrying fluid phase, have now to be considered as potentially affecting flow regimes. It shall be noted that in some specific situations, some of these dimensionless parameters could be linked (for instance $\phi$ and $\mathcal{A}t$ for heavy particles in a single-phase liquid for the carrying and ambient fluids).

>>>>

Other mechanisms can modify the dynamics of the slumping regime, possibly leading to other possible positions $X_{\text t}$ or times $T_{\text t}$ at which the velocity is no longer time-independent. Depending on the relative intensity of these mechanisms, this may occur prior to the influence of the rarefaction wave or bore. Hence, determining the various regimes related to the prevalence of each mechanism is part of the topic of the present chapter.

@@

A major difference between PLGCs and compositional currents is therefore the existence of typical length and time scales associated with the particle dynamics, such as settling or particle/particle interactions. Nevertheless, the prediction of these finite scales is difficult as it should strongly depend on the interaction between particle dynamics and the local turbulence of the carrying fluid and/or ambient fluid at the upper interface which modifies entrainment and associated dissipation. In turn, this can modify the macroscopic behavior of the current, for instance, the dynamics of the front. Accordingly, these dynamics shall be controlled by several dimensionless parameters that can lead to a wide variety of flow regimes. In particular, beyond the slope $\alpha$, the dimensionless density difference between the current and the ambient $\mathcal{A}t$, the ratio between available inertia and dissipation, say a Reynolds number $\mathcal{R}e$ as usually introduced as a control parameter for any density currents, the initial solid fraction of particles $\phi$ and a dimensionless settling velocity $S=V_{\text s}/U_{\text{c}}$ with $V_{\text s}$ the settling velocity of particles into the carrying fluid phase, have now to be considered as potentially affecting flow regimes. It shall be noted that in some specific situations, some of these dimensionless parameters could be linked (for instance $\phi$ and $\mathcal{A}t$ for heavy particles in a single-phase liquid for the carrying and ambient fluids).
